---
title: "Plot and Analyze microbial plate growth data"
author: "Maximilian Berthold, Douglas A. Campbell"
date: "`r format(Sys.Date())`"
output:
  html_document:
    df_print: paged
---

Set variables for project
```{r project data}
Project <- "PGW"
FileIDSingle <- "Chl" #possible factors: Chl OD CC
FP1 <- ".."
FP2 <- "TidiedData"


#grouping variables for rows & columns, and for nesting for fitting
RowVar <- "temp_c" #expected variable, like temperature, salinity
#RowVar2 <- "CalcSal" #possible second grouping variable
ColVar <- "strain" #always group by strain
NestVar <- "par_ue" #group by light level, if applicable 
NestVarII <- "id" #group by culture ID to have each well individually fitted, alternatively this can be 'well'

#filtering variables
nm_filter <- c("680min750") #"680min720", "680min750", "720", "750" different wavelengths important to estimate culture growth; 680min720 and 680 min 750 represent deltaOD of those wavelengths to get a clear Chlorophyll signal; OD720 and 750 are wavelengths used for scattering, to determine ~cell numbers


StartDate <- c("20210825") #if only a selection of days should be fitted at once
StopTime = 10 #time in days, after which fit should be stopped

#fitting variable
FitVar <- "RegresChl_ugL" #CorrOD RegresCellCount_mL RegresChl_ugL


```



```{r load libraries}
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(knitr)
library(minpack.lm)
library(nlstools)
library(nlsMicrobio)

```



```{r read data, warning=FALSE}
ImportGrowth <- readRDS(file = file.path(paste(FP1, FP2, sep = "/"),paste(Project,FileIDSingle, "GrowthLong.Rds", sep  = "_"),fsep = .Platform$file.sep))

#https://stackoverflow.com/questions/64111558/passing-variables-into-the-names-glue-parameter-of-tidyrpivot-wider

GrowthLong <- ImportGrowth %>%
  filter(Wavelength %in% nm_filter) #%>%
  # pivot_wider(id_cols = -c(OD, AvBlankOD), names_from = Wavelength, values_from = CorrOD, names_glue = "nm_{Wavelength}") %>% 
  # mutate("680min720" = nm_680 - nm_720, "680min750" = nm_680 - nm_750) %>%
  # pivot_longer(cols = c(nm_680, nm_720, nm_750, "680min720", "680min750"), values_to = "CorrOD", names_to = "Wavelength") %>%
  # mutate(Wavelength = str_remove(Wavelength, "nm_")) %>%
  # mutate(CorrOD = if_else(CorrOD < 0.001, 0.001, CorrOD)) #experimentally test to set values below 0.001 to 0.001 (~detection limit device)


GrowthLong <- GrowthLong %>%
  group_by(!!sym(NestVar), !!sym(RowVar), !!sym(ColVar), !!sym(NestVarII), exp_date) %>%
  mutate(logFitVarminFitVar = log(!!sym(FitVar)/min(!!sym(FitVar), na.rm = TRUE))) %>% #problem with log leads to NaN, when value is very small
  ungroup() %>%
  mutate(strain = as.factor(strain))


```



```{r noise filtering}
#set filtering variables at top
#pick filtering option as either a simple threshold or a factor change between OD and OD_lead

#Screen_OD <- 2

FitData <- as_tibble(GrowthLong) %>%
  filter(exp_date %in% StartDate) %>% #if multiple dates are applied, use %in% instead of ==
  #filter(Wavelength %in% Wavelength_nm) %>%
  #filter(tissue_treated_plate == TissueTreated) %>%
  #filter(media == Media) %>%
  #filter(source_salinity == SourceSal) %>%
  group_by(well) %>%
  arrange(well, E_days) %>%
  mutate(FitVar_lead = lead(!!sym(FitVar), na.rm = TRUE)) %>%
  #filter(CorrOD < (OD_lead / Screen_OD)) %>%
  #filter(CorrOD < Screen_OD) %>%
  filter(E_days < StopTime)

test <- FitData %>%
  group_by(!!sym(NestVar), !!sym(RowVar), exp_date) %>%
  mutate(LagSeed = E_days[which.min(!!sym(FitVar))])
  
```


```{r filter plot}
FitData %>%
  filter(exp_date %in% StartDate) %>%
  #filter(Wavelength %in% Wavelength_nm) %>%
  #filter(tissue_treated_plate == TissueTreated) %>%
  #filter(media == Media) %>%
  ggplot()+
  geom_point(aes(x = E_days, y = !!sym(FitVar), colour = as.factor(salinity))) +
  #coord_cartesian(ylim = c(0, 2)) +
  facet_grid(cols = vars(exp_date), rows = vars(!!sym(NestVar))) +
    theme_bw() #+
  #labs(caption = paste(Wavelength_nm, "nm; PAR", Par_ue, "uE"))
```


Define equations as functions.
x will be taken from 'E_days' when we run the fit.
```{r logistic_eqn}
LogisticEqn <-  function(x, Pmax, Mu, Intercept){(Pmax*Intercept*exp(Mu*x))/(Pmax + (Intercept*(exp(Mu*x)-1)))
}

ModGompertzEqn <- function(x, Gmax,Gmu,Lag){(Gmax*(exp(-exp((Gmu*exp(1))/Gmax*(Lag-x)+1))))}

possibnlSLM = possibly(.f = nlsLM, otherwise = NULL)
```


Fit treatment specific logistic growth trajectories using nest purrr:map & broom::augment using nest specific start, lower & upper settings extracted from each nest on the fly.
This may be necessary if the 'nests' contain diverse data patterns that fail to fit with generic start, lower & upper parameters extracted from the entire dataset.
```{r treat growth specific}
#run nesting by passing in values of variables set at top; 
#!!sym(... allows nest to access the characters of the values passed in
TreatNestSpecific <- test %>%
  nest(data = -c(!!sym(NestVar), !!sym(RowVar), !!sym(RowVar), !!sym(ColVar), Wavelength, LagSeed, exp_date))


#extract nest-specific start, lower & upper setting by passing in "."
TreatLogSpecific <- TreatNestSpecific %>% 
    mutate(
  FitLog = map(data, ~possibnlSLM(!!sym(FitVar) ~ LogisticEqn(x = E_days, Intercept, Mu, Pmax),
                            data = .x, 
                            start = list(
                              Intercept = min(.[,FitVar], na.rm = TRUE), 
                              Mu = (log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days), 
                              Pmax = max(.[,FitVar], na.rm = TRUE)),
                            control = list(maxiter = 10000)
                            ,
                            lower = c((min(.[,FitVar], na.rm = TRUE) * 0.01),
                                      ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days)) *0.01,
                                      ((max(.[,FitVar], na.rm = TRUE)) * 0.01)),
                            upper = c((min(.[,FitVar], na.rm = TRUE) * 10),
                                      ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days)) * 10,
                                      ((max(.[,FitVar], na.rm = TRUE)) * 10))
               )
                ),
  PredictLog = map(FitLog, augment),
  TidiedLog = map(FitLog, tidy),
  ParamLog = map(FitLog, glance)
  ) %>%
   mutate(
    FitGompertz = (map(data, ~nlsLM(logFitVarminFitVar ~ ModGompertzEqn(x = E_days, Lag, Gmu, Gmax),
                                    data = .x,
                                    start = list(Lag = LagSeed, Gmu = (log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days), Gmax = max(.[, "logFitVarminFitVar"], na.rm = TRUE)),
                                    lower = c(0, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days)) * 0.01, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 0.01),
                                    upper = c(48, ((log(max(.[, FitVar], na.rm = TRUE)) - log(min(.[ FitVar], na.rm = TRUE)))/max(.$E_days)) * 1.5, max(.[, "logFitVarminFitVar"], na.rm = TRUE) * 1.5)))),
    PredictGompertz = map(FitGompertz, augment),
    TidiedGompertz = map(FitGompertz, tidy),
    ParamGompertz = map (FitGompertz, glance))

TreatLogSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictLog)

TreatLogSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedLog) %>%
  dplyr::select(c(!!sym(NestVar), !!sym(RowVar2), !!sym(RowVar), !!sym(ColVar), Wavelength, term:p.value))

TreatGompertzSpecificPredict <- TreatLogSpecific %>%
  unnest(PredictGompertz)

TreatGompertzSpecificCoeff <- TreatLogSpecific %>% 
  unnest(TidiedGompertz) %>%
  dplyr::select(c(!!sym(NestVar), !!sym(RowVar2), !!sym(RowVar), !!sym(ColVar), Wavelength, term:p.value))
```


```{r}
CheckDate <- StartDate %>%
  as_data_frame(StartDate) %>%
  slice(1)

TreatLogSpecificPlot <- TreatLogSpecificPredict %>% 
  filter(exp_date %in% CheckDate) %>%
  filter(Wavelength %in% nm_filter) %>%
  ggplot() +  
  geom_line(aes(x = E_days, y = .fitted), size = 0.5) +
  geom_point(aes(x = E_days, y = !!sym(FitVar)), size = 0.6) +
  # geom_ribbon(aes(x = E_days, ymin = (.fitted - .resid), ymax = (.fitted + .resid), alpha = 0.1),show.legend = FALSE) +
  facet_grid(cols = vars(CalcSal), rows = vars(par_ue)) +
  theme_bw() +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text = element_text(colour = "black")) +
  #labs(y = expression(paste("log(OD"[`680 nm`]*"/min OD"[`680 nm`]*")")), x = "Elapsed time (h)", title = "Treatment Growth, Modified Gompertz with Lag", subtitle = paste(str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_")))
  labs(y = FitVar,
       title = "Treatment Growth, Specific Log",
       subtitle = paste(nm_filter, "nm", str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_")),
       caption = "Log")

TreatLogSpecificPlot

TreatGompertzLagSpecificPlot <- TreatGompertzSpecificPredict %>% 
  filter(exp_date %in% CheckDate) %>%
  filter(Wavelength %in% nm_filter) %>%
  ggplot() +  
  geom_line(aes(x = E_days, y = .fitted), size = 0.5) +
  geom_point(aes(x = E_days, y = logFitVarminFitVar), size = 0.6) +
  # geom_ribbon(aes(x = E_days, ymin = (.fitted - .resid), ymax = (.fitted + .resid), alpha = 0.1),show.legend = FALSE) +
  facet_grid(cols = vars(CalcSal), rows = vars(par_ue)) +
  theme_bw() +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text = element_text(colour = "black")) +
  #labs(y = expression(paste("log(OD"[`680 nm`]*"/min OD"[`680 nm`]*")")), x = "Elapsed time (h)", title = "Treatment Growth, Modified Gompertz with Lag", subtitle = paste(str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_")))
  labs(y = FitVar,
       title = "Treatment Growth, Modified Gompertz Specific with Lag",
       subtitle = paste(nm_filter, "nm", str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_")),
       caption = "Modified Gompertz with Lag Line")

TreatGompertzLagSpecificPlot
```


```{r save plots}
# ggsave(filename = file.path("Plots", paste(Project, Replic, 
# FileIDSingle, str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_"),"TreatLogPlot.png", sep = "_"),fsep = .Platform$file.sep ), plot = TreatLogPlot, height = 4, width = 6)
# 
# ggsave(filename = file.path("Plots", paste(Project, Replic, 
# FileIDSingle, str_c(Strains, collapse = "_"), str_c(StartDate, collapse = "_"), "TreatGompertzLagPlot.png", sep = "_"),fsep = .Platform$file.sep ), plot = TreatGompertzLagPlot, height = 4, width = 6)
```

Display & save parameters of logistic & modified Gompertz with lag.
```{r fit parameters}
TreatLogSpecificCoeff %>%
  mutate_if(is.numeric, round, digits = 3)

TreatGompertzSpecificCoeff %>%
  mutate_if(is.numeric, round, digits = 3)

#save .rds of the current round of fit parameters
saveRDS(TreatLogSpecificCoeff, file = file.path(DataFolder,paste(Project,FileIDSingle, Replic, "TreatLogSpecificCoeff", str_c(Strains, collapse = "_"), "ChlAb", ".Rds",sep = "_"),fsep = .Platform$file.sep))

#, str_c(StartDate, collapse = "_") add, if not all replicates are run at once

saveRDS(TreatGompertzSpecificCoeff, file = file.path(DataFolder,paste(Project,FileIDSingle, Replic, "TreatGompSpeCo", str_c(Strains, collapse = "_"), "ChlAb", ".Rds",sep = "_"),fsep = .Platform$file.sep))

```

